---
title: "DATA 607 Web APIs"
author: "Josh Iden"
date: "2022-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task: choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

I'm going to use the jsonlite package to read in the JSON and transform it to a dataframe.

```{r warning=FALSE, message=FALSE}
library(jsonlite)
```

Storing the apiKey (hidden) and url:

```{r echo=FALSE}
apiKey <- "GJ6QegUp6KrW2ETMXAd9BqwKj3h32hGZ"
```

```{r}
url <- "https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key="
api <- paste(url,apiKey,sep="")
```

Using fromJSON, I read in the data to latest_fiction, and found the book data in the nested list books inside results:

```{r}
latest_fiction <- fromJSON(api) 
df <- as.data.frame(latest_fiction$results$books)
head(df)
```

names() in conjunction with view() gives me a starting point to tidy my dataframe for future analysis.

```{r}
names(df)
```

Previewing a subset of the dataframe:

```{r}
cols = c(11,12,1:3)
head(df[cols])
```
