---
title: "DATA607_HW7"
author: "Josh Iden"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For this assignment, I picked three books from one of my favorite genres: horror fiction, and created three files in which I stored each book's information, and read each file into R:

[HTML](https://github.com/josh1den/DATA-607/blob/main/HW/HW7/DATA607_HW7.html)

[XML](https://github.com/josh1den/DATA-607/blob/main/HW/HW7/DATA607_HW7.xml)

[JSON](https://github.com/josh1den/DATA-607/blob/main/HW/HW7/DATA607_HW7.json)

## Reading Into R

For this assignment, I utilized the tidyverse package in order to pipe my data into R for readability. The individual packages I will note with their corresponding file formats for comprehension.

```{r}
library(tidyverse)
```

## Reading HTML

```{r}
library(htmltab)

file <- "https://raw.githubusercontent.com/josh1den/DATA-607/main/HW/HW7/DATA607_HW7.html"
html <- htmltab(file, which = 1) |>
        as.data.frame()
print(html)
```

## Reading XML

```{r}
library(XML)
library(xml2)
file <- "https://raw.githubusercontent.com/josh1den/DATA-607/main/HW/HW7/DATA607_HW7.xml"

xml <- xml2::read_xml(file) |>
       XML::xmlParse() |>
       XML::xmlToDataFrame()

print(xml)
```

## Reading JSON

```{r}
library(rjson)
file <- "https://raw.githubusercontent.com/josh1den/DATA-607/main/HW/HW7/DATA607_HW7.json"
json <- fromJSON(file=file) |>
        as.data.frame()
print(json)
```
While the HTML and XML dataframes each approximate one another, the JSON does not. 
Upon investigation, I discovered that the format of my JSON was providing challenges I was unable to resolve. By [altering my JSON format](https://github.com/josh1den/DATA-607/blob/main/HW/HW7/DATA607_HW7_V2.json), I was able to read in to R delivering an output equivalent to those of the XML and HTML formats:

```{r}
file_v2 <- "https://raw.githubusercontent.com/josh1den/DATA-607/main/HW/HW7/DATA607_HW7_V2.json"
json_v2 <- fromJSON(file=file_v2) |>
           as.data.frame()
print(json_v2)
```

While I did not resolve the output challenge for the first version JSON file, one insight I glean from this is that JSON formatting can have a major effect on its output, and understanding the structure of your source file is essential to crafting code to achieve the desired output. 
