---
title: "DATA607_HW3"
author: "Josh Iden"
date: "2022-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message=FALSE}
# load packages
library(rvest)
library(htmltab)
library(stringr)
library(dplyr)
```

--

#### 1.Using the 173 majors listed in fivethirtyeight.com's College Majors dataset [<https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/>], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r 1}
# store file URL
file <- "https://projects.fivethirtyeight.com/mid-levels/college-majors/index.html"
data <- read_html(file)

# read the table from the data
allTables <- data %>%
  html_table(fill=TRUE) 

# select the first table as a tibble
table1 <- allTables[[1]]
# create a vector of the majors
majors <- collect(select(table1, MAJOR))[[1]]
# detect matches
matches <- str_detect(majors, "(?i)statistics|(?i)data")
# subset majors
majors[matches]
```

--  

#### 2. Write code that transforms the data below:

[1] "bell pepper" "bilberry" "blackberry" "blood orange"\
[5] "blueberry" "cantaloupe" "chili pepper" "cloudberry"\
[9] "elderberry" "lime" "lychee" "mulberry"\
[13] "olive" "salal berry"

Into a format like this:\
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r 2}
text <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

# stringr extract_all
words <- str_extract_all(text, regex("\\w+..\\w+"))
# simplify list to a character vector
unlist(words)
```
--  

#### 3. Describe, in words, what these expressions will match:

-   (.)\\1\\1\
    *anything appearing three times in a row*

-   "(.)(.)\\2\\1"\
    *two things followed by those same two things in reverse*

-   (..)\\1\
    *any two things repeated once*

-   "(.).\\1.\\1"\
    *anything followed by anything followed by the first thing followed by anything followed by the first thing again*

-   "(.)(.)(.).*\\3\\2\\1"\
    *any three things followed by 0 or more characters of anything followed by the first three things in reverse\*

--  

#### 4. Construct regular expressions to match words that:

-   Start and end with the same character: *"(.).+\\1\$"*

```{r 4.1}
same <- c("yay", "bcb", "dad", "roger", "maaaaan", "maaaaaaaaaam")
str_view(same, "(.).+\\1$", match=TRUE)
```

-   Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.): *"(..).+\\1"*

```{r 4.2}
church <- c("church", "barn", "outhouse", "papa", "eleven")
str_view(church, "(..).+\\1", match=TRUE)
```

-   Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.): *"([a-zA-Z]).\\1.+\\1.+"*

```{r 4.3}
str_view(church, "([a-zA-Z]).\\1.+\\1.+", match=TRUE)
```
