---
title: "DATA607 HW11"
author: "Josh Iden"
date: "2022-11-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Introduction

In this assignment, I'm tasked to apply some sentiment analysis base code from chapter 2 in [Text Mining With R](https://www.tidytextmining.com/) and apply it to a corpus of my choosing.

#### The Code

First let's take a look at the code from chapter 2. The code features the following package dependencies:

```{r packages messages=FALSE}
library(tidytext)
library(tidyverse)
library(janeaustenr)
```
About the packages:

**tidytext** provides functions for tidying text
**janeaustenr** is a sample dataset of Jane Austen's 6 novels for text analysis. 

##### Loading Data

First we'll load joy words from the [NRC lexicon](https://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm) and the Jane Austen dataset, which we'll also tidy.

```{r}
#load nrc lexicon, filtering for joy
nrcjoy = tidytext::get_sentiments("nrc") |>
  filter(sentiment == "joy")
# load and tidy the jane austen data
tidy_books = austen_books() |>
  group_by(book) |>
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", ignore_case=TRUE)))) |>
  ungroup() |>
  tidytext::unnest_tokens(word, text)
```

Let's take a look at the output files:

```{r}
head(nrcjoy)
```
```{r}
head(tidy_books)
```
##### Calculating Sentiment

Now we'll combine the nrc sentiment data with the Jane Austin data to get a count of joy words that appear in the novel "Emma". 

```{r}
tidy_books |>
  filter(book == "Emma") |>
  inner_join(nrcjoy) |>
  count(word, sort=TRUE)
```

We can also use the [BING](https://emilhvitfeldt.github.io/textdata/reference/lexicon_bing.html) lexicon which assigns a binary value to each word (positive or negative) to calculate a net sentiment.

```{r}
# calculate a net sentiment
janeaustensentiment = tidy_books |>
  inner_join(get_sentiments("bing")) |>
  count(book, index = linenumber %/% 80, sentiment) |>
  spread(sentiment, n, fill = 0) |>
  mutate(sentiment = positive - negative)

head(janeaustensentiment)
```

#### A New Corpus
